# Data sources
---
title: "Data Sources"
author: "Rahul Subramaniam, Santos Hernandez"
output:
  html_document: default
  pdf_document: default
---

```{r, include=TRUE}
SOLUTIONS <- FALSE
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = SOLUTIONS, 
                      eval = SOLUTIONS, cache = FALSE)
library(tidyverse)
```


Question :

In the NBA, is injury related to position or player statistics in some way? Has this changed over time through certain eras of playstyle?

```{r}
# keep this chunk in your .Rmd file
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


The NBA tools are available as a part of the nbaStatR package. We install it using the devtools package offered by github.


```{r}
devtools::install_github("abresler/nbastatR",force = TRUE)
```

Showing the libraries

```{r}
library(nbastatR)
library(future)
plan(multiprocess) 
```



c) Sampole view of Data 
```{r}
game_logs(seasons = 2010:2019)
```


```{r}
bref_injuries()
```

b) Read the page at the URL listed above and save it.

c) Read in the page saved in part b) and create a data frame with the following columns: `State`, `Distributed`, `Administered`, `Percentage`. Display the first six rows of the data frame.

d) Create a Cleveland dot plot showing the percentage of distributed vaccines that have been administered.
```{r}

#players_rotowire(players = c( "Jarrett Allen", "DeMarre Carroll", "Allen Crabbe"))
playerData<-dictionary_bref_players()


```

e) Create a multiple Cleveland dot plot showing doses distributed in one color and doses administered in another for each state. Experiment with different sorting orders. Choose the one that you think works best and explain why.

```{r}
library(tibble)
c<-bref_bios( players = c("De'Andre Hunter", "Mitch Richmond", "Michael Adams"),
player_ids = NULL,
assign_to_environment = TRUE)['dataTable']
c
```



```{r}
players_careers(players = c("Joe Harris", "Myles Turner", "Spencer Dinwiddie"),
modes = c("Totals", "PerGame"))
```


### 3. Taxis

Data: NYC yellow cab rides in June 2020, available here:  

https://www1.nyc.gov/site/tlc/about/tlc-trip-record-data.page

It's a large file so you may work with a reasonably-sized random sample of the data (for example 10000 rows). If you remove outliers / impossibilities explain your logic.

Draw four plots of `tip_amount` vs. `fare_amount` with the following variations:

(a) Scatterplot with alpha blending

(b) Scatterplot with alpha blending + density estimate contour lines

(c) Hexagonal heatmap of bin counts

(d) Square heatmap of bin counts 

For all, adjust parameters to the levels that provide the best views of the data.

(e) Describe noteworthy features of the data, using the "Movie ratings" example on page 82 (last page of Section 5.3) as a guide.  


## Game statistics
In the NBA, the largest source of game statistics comes from the league itself. This data is 
publicly available at the [NBA website](https://www.nba.com/stats/), with some stats going back as far as 1946.
Scorekeepers are hired to watch every single game, tracking every statistic that is available to us. Many of the 
statistics are raw counts of things like points, minutes played, and rebounds, while others are 
calculated metrics. This does not discount the fact that the stats we see are 
subject to [human error](https://deadspin.com/the-confessions-of-an-nba-scorekeeper-5345287),
but generally, the NBA is the best data source available for game statistics. 

[NBAstatR](https://github.com/abresler/nbastatR) is a package that crawls the website to return the data directly in R. The game data is aggregated 
and rolled up based on specified parameters. For example, one could request the per-game averages 
of each team for the 2015 season, from March to April. The package would return a table 
with a row for each team, with average statistics for allgames played in that time period. This reflects the behavior of the NBA stats 
website, which does the same thing. There are 30 teams in the modern NBA, so 30 rows are always returned in requests for team statistics. 
When requesting player data, the number of rows can vary because not every player plays every game and people are constantly getting added/dropped from rosters. 
Generally, for the whole regular season, there are between 500-550 rows of unique player data. Less are returned for 
requests of smaller windows of time, as some players may not play at all during that time. 

Basketball stats are complicated. There are a seemingly endless amount of statistics 
that are available for each player and team. For example, the "Traditional" stats reported include things 
like 3 Point Field Goals Made, Free Throw Percentage, Rebounds, Assists, and Steals, among many others. 
For our project, we were most interested understanding _general_ team/player performance, which can be 
difficult to discern from all of those statistics. So, we focused on the 
Offensive, Defensive, and Net ratings. The Offensive rating can be understood as 
the amount of points produced on offense when the player was on the court, while Defensive rating can be understood as 
the amount of points allowed on defense when the player was on the court. The way these are 
calculated are somewhat [complicated](https://www.basketball-reference.com/about/ratings.html), 
but are a good way to understand general performance. Team versions of these ratings are calculated 
based on the underlying player statistics.

### COVID cases

A single table was scraped from a [basketball magazine's website](https://www.slamonline.com/nba/nba-coronavirus-covid-list/) 
(SLAM) to get a list of players that were confirmed to have COVID, and when it was reported by the magazine.

## Player positions and injuries

We also needed other data that went beyond game statistics: player position data and injury data. 
The NBA also makes this data somewhat available on its website, does not aggregate and release it 
as conveniently. Player position data is limited only to forward, guard, and center, 
while most think of the game as played in 5 positions - point guard, shooting guard, small forward, power forward, and center. 
Injury data available on the website is only for current injuries, and is not historical. 

////EXPAND VVV
We turned to [Basketball Reference](https://www.basketball-reference.com/about/ratings.html) for more complete player profiles. T
his data is collected by a parent organization that manages 
several sports databases. The core of its historical data [comes from](https://www.basketball-reference.com/about/sources.html) 
Sean Lahman, a researcher/journalist famous for compiling open source sports data and encyclopedias.

Injury data was scraped from [Pro Sports Transactions](http://www.prosportstransactions.com/), which compiles data that is released by the NBA.

## Issues

When we compare a "month" of game data for each team, each team may not have played 
the same amount of games due to scheduling. The sample for each average can be 
slightly unequal by +/- a game played from team to team. The solution to this would be to obtain averages based 
on the number of games played before a certain date, but the API lacks this slicing capability. 
It might be possible pull the box scores of every single game, count out and filter based on date, and then create the averages manually, 
but this approach would be a significant amount of work for perhaps a small change in results. As previously stated, 
we are looking for differences in general performance.
